---
title: 'Marginal means with lmer'
output: pdf_document
classoption: landscape
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y %h:%m')`"
---

# Summary
Just confirming that emmeans does what I think it does with mixed-effects models without balance.

# 0. Packages

```{r}
library(lme4)
library(emmeans)
library(data.table)
```

```{r}
source('dfcombos.R')
```



# 1. Data

Five levels of a random effect.

```{r}
rv <- data.table(rv = factor(letters[1:5]), re = -2:2)
rv
```

(Hmm, should probably try with a mean different from 0.)

Three treatments make up a fixed effect.

```{r}
ff <- data.table(trt = factor(letters[1:3]), fe = c(-1, 0, 1))
```

Combine.

```{r}
d0 <- dfcombos(rv, ff)
```

Eliminate balance.

```{r}
d0 <- d0[-c(9:10, 12:14), ]
d0
table(d0$trt)
table(d0$rv)
```

Add response variable

```{r}
setDT(d0)
d0[, y := re + fe + rnorm(nrow(d0), sd = 0.2)]
d0
```

# 2. Model

```{r}
m1 <- lmer(y ~ trt + (1|rv), data = d0)
summary(m1)
emmeans(m1, 'trt')
```

Compare to mean by treatment.

```{r}
d0[, mean(y), by = trt]
```

# 3. Higher n

Instead of running this many times, I'll just increase the sample size.

```{r}
d1 <- d0[rep(1:nrow(d0), each = 300), ]
table(d1$trt)
```

And recreate `y` values.

```{r}
set.seed(1)
d1[, y := re + fe + rnorm(nrow(d1), sd = 0.2)]
d1
```

```{r}
m2 <- lmer(y ~ trt + (1|rv), data = d1)
summary(m2)
emmeans(m2, 'trt')
```

Compare to mean by treatment.

```{r}
d1[, mean(y), by = trt]
```

Notice that `lmer` coefficients (`Estimate`) and marginal means look identical.

# Conclusion
The emmeans package seems to be a simple way to get marginal means from mixed-effects models.


















